---
title: "Assignment 1"
description: |
  VAST Challenge 2021
author:
  - name: Davmes Tan
    url: https://www.linkedin.com/in/davmestan/
date: 07-13-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)
```

```{r packages}

packages = c('raster', 'sf',
             'clock', 'tidyverse', 
             'tmap', 'rgdal',
             'dplyr', 'ggplot2',
             'ggiraph', 'DT',
             'plotly', 'crosstalk',
             'forcats', 'scales')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)

}

```

# 1. Background

## 1.1 Background Context

Extracted from VAST Challenge 2021 [here](https://vast-challenge.github.io/2021/index.html)

*In the roughly twenty years that Tethys-based GAStech has been operating a natural gas production site in the island country of Kronos, it has produced remarkable profits and developed strong relationships with the government of Kronos. However, GAStech has not been as successful in demonstrating environmental stewardship.*

*In January, 2014, the leaders of GAStech are celebrating their new-found fortune as a result of the initial public offering of their very successful company. In the midst of this celebration, several employees of GAStech go missing. An organization known as the Protectors of Kronos (POK) is suspected in the disappearance, but things may not be what they seem.*

## 1.2 The Challenge

Extracted from VAST Challenge 2021 Mini Challenge 2 [here](https://vast-challenge.github.io/2021/MC2.html).

*Many of the Abila, Kronos-based employees of GAStech have company cars which are approved for both personal and business use. Those who do not have company cars have the ability to check out company trucks for business use, but these trucks cannot be used for personal business.*

*Employees with company cars are happy to have these vehicles, because the company cars are generally much higher quality than the cars they would be able to afford otherwise. However, GAStech does not trust their employees. Without the employees' knowledge, GAStech has installed geospatial tracking software in the company vehicles. The vehicles are tracked periodically as long as they are moving.*

*This vehicle tracking data has been made available to law enforcement to support their investigation. Unfortunately, data is not available for the day the GAStech employees went missing. Data is only available for the two weeks prior to the disappearance.*

*To promote local businesses, Kronos based companies provide a Kronos Kares benefit card to GASTech employees giving them discounts and rewards in exchange for collecting information about their credit card purchases and preferences as recorded on loyalty cards. This data has been made available to investigators in the hopes that it can help resolve the situation. However, Kronos Kares does not collect personal information beyond purchases.*

The objective of this assignment is to assist law enforcement authorities to **ascertain varying purchases made by specific GASTech employees** and to **identify suspicious patterns of behaviour**.

# 2. Data Preparation

There are a total of 3 csv files provided for MC 2. They are:

+ **cc_data.csv**
+ **loyalty_data.csv**
+ **gps.csv**

```{r data_prep_1}

cc_data <- read_csv("data/cc_data.csv")
loyalty_data <- read_csv("data/loyalty_data.csv")

#convert timestamp from character into date/time format
cc_data$timestamp <- date_time_parse(cc_data$timestamp,
                                     zone = "",
                                     format = "%m/%d/%Y %H:%M")

#convert timestamp from character into date/time format
loyalty_data$timestamp <- date_time_parse(loyalty_data$timestamp,
                                     zone = "",
                                     format = "%m/%d/%Y")

#Amend string text for Katrina's Cafe
cc_data2 <- cc_data %>%
  mutate(location = str_replace_all(location,
                                    pattern = "Katerin.+",
                                    replacement = "Katrina\x27s Caf\xe9"))
cc_data <- cc_data2

loyalty_data2 <- loyalty_data %>%
  mutate(location = str_replace_all(location,
                                    pattern = "Katerin.+",
                                    replacement = "Katrina\x27s Caf\xe9"))
loyalty_data <- loyalty_data2

```



```{r Question1a}

cc_plot <- ggplot(cc_data) + 
  geom_histogram_interactive(aes(fct_infreq(location),
                                 stackgroups = TRUE,
                                 binwidth = 1),
                             fill = "orange",
                             stat = "count") +
  geom_text(aes(location, label=after_stat(count)),
            stat = 'count',
            position = position_dodge(width = 0.9), 
            size=2.5, 
            vjust = -1) + 
  ylim(0,230) + 
  labs(x = "Locations", y = "Number of Txn", title = "Visits based on Credit Card Transactions") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        plot.title = element_text(hjust = 0.5))

loyalty_plot <- ggplot(loyalty_data) + 
  geom_histogram_interactive(aes(fct_infreq(location),
                                 stackgroups = TRUE,
                                 binwidth = 1),
                             fill = "blue",
                             stat = "count") +
  geom_text(aes(location, label=after_stat(count)),
            stat = 'count',
            position = position_dodge(width = 0.9), 
            size=2.5, 
            vjust = -1) + 
  ylim(0,230) + 
  labs(x = "Locations", y = "Number of Txn", title = "Visits based on Loyalty Card Transactions") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        plot.title = element_text(hjust = 0.5))

cc_plot
loyalty_plot
```

It was observed that the following 4 locations were highly patronised regardless whether it is a credit card transaction or a loyalty card transaction:

+ Katrina's Café
+ Hippokampos
+ Guy's Gyros
+ Brew've Been Served

```{r Question1b}

sel_cc_data <- cc_data %>%
  filter(location == "Katrina's Café" |
           location == "Hippokampos" |
           location == "Guy's Gyros" |
           location == "Brew've Been Served")

sel_cc_data <- sel_cc_data %>%
  mutate(day = get_day(timestamp),
         weekday = weekdays(timestamp),
         hour = get_hour(timestamp),
         time = format(timestamp, format = "%H:%M:%S"))

sel_cc_data$time <- date_time_parse(sel_cc_data$time,
                                     zone = "",
                                     format = "%H:%M:%S")

#########################################################
#Setting Y axis as time
#########################################################
ggplot(sel_cc_data, aes(x = location, y = time,
                        group = weekday)) + 
  geom_boxplot()+
  scale_y_datetime(labels = function(x) format(x, "%H:%M", tz = "EST"),
                   date_breaks = '1 hour',expand = c(0,0),
                   limits = as.POSIXct(c('00:00:00','24:00:00'),
                                       format="%H:%M:%S"),
                   date_labels = c('00:00','00:01','00:02','00:03',
                                   '00:04','00:05','00:06','00:07',
                                   '00:08','00:09','10:00','11:00',
                                   '12:00','13:00','14:00','15:00',
                                   '16:00','17:00','18:00','19:00',
                                   '20:00','21:00','22:00','23:00',
                                   '24:00'))+
  xlab('Location')


```
