---
title: "Assignment 1"
description: |
  VAST Challenge 2021
author:
  - name: Davmes Tan
    url: https://www.linkedin.com/in/davmestan/
date: 07-13-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)
```

```{r packages}

packages = c('raster', 'sf',
             'clock', 'tidyverse', 
             'tmap', 'rgdal',
             'dplyr', 'ggplot2',
             'ggiraph', 'DT',
             'plotly', 'crosstalk')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)

}

```

```{r old q1 plot}

cc_plot <- ggplot(cc_data) + 
  geom_histogram_interactive(aes(fct_infreq(location),
                                 stackgroups = TRUE,
                                 binwidth = 1),
                             fill = "orange",
                             stat = "count") +
  geom_text(aes(location, label=after_stat(count)),
            stat = 'count',
            position = position_dodge(width = 0.9), 
            size=2.5, 
            vjust = -1) + 
  ylim(0,230) + 
  labs(x = "Locations", y = "Number of Txn", title = "Visits based on Credit Card Transactions") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        plot.title = element_text(hjust = 0.5))

loyalty_plot <- ggplot(loyalty_data) + 
  geom_histogram_interactive(aes(fct_infreq(location),
                                 stackgroups = TRUE,
                                 binwidth = 1),
                             fill = "blue",
                             stat = "count") +
  geom_text(aes(location, label=after_stat(count)),
            stat = 'count',
            position = position_dodge(width = 0.9), 
            size=2.5, 
            vjust = -1) + 
  ylim(0,230) + 
  labs(x = "Locations", y = "Number of Txn", title = "Visits based on Loyalty Card Transactions") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        plot.title = element_text(hjust = 0.5))

time_cc_data <- cc_data %>%
  mutate(time = format(timestamp, format = "%H:%M"))

#time_cc_data$time <- date_time_parse(time_cc_data$time,
#                                     zone = "",
#                                     format = "%H:%M")

for(i in 5:18){
  
  st <- paste("01/", i , "/2014 16:00", sep ="")
  et <- paste("01/", i+1 , "/2014 15:59", sep ="")
  
  start_time <- mdy_hm(st)
  end_time <- mdy_hm(et)
  
  time_cc_data <- cc_data %>%
  filter(timestamp > start_time &
           timestamp <end_time)

  time_cc_data <- time_cc_data %>%
    mutate(time15 = round_date(time_cc_data$timestamp, "15 minutes"))
  
  time_cc_data <- time_cc_data %>%
    group_by(location, time15) %>%
    add_count(location, time15, name = "count")
  
  pop_time_cc_data <- time_cc_data %>%
    filter(location == "Katrina's Café"|
             location == "Hippokampos" |
             location == "Guy's Gyros" |
             location == "Brew've Been Served")
  
  gTitle <- paste("Transaction Counts on ", i+1, "/01/2014", sep="")
  
  cc_tm_plot <- ggplot(pop_time_cc_data,
         aes(time15, count, group = location)) + 
    geom_line(aes(color = location), size = 1.5) + 
    geom_point() + 
    scale_fill_brewer(palette = "Set2") +
    labs(x = "Time", y = "Number of Txn", title = gTitle) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "top")
  
  svl <- paste("Top 4 on ", i+1, "-01-2014.png", sep="")
  #ggsave(svl, cc_tm_plot)
  
}
```

```{r Weekday Weekends Visits}

sel_cc_data <-cc_data %>%
  filter(location == "Katrina's Café" |
           location == "Hippokampos" |
           location == "Guy's Gyros" |
           location == "Brew've Been Served")

sel_cc_data <- sel_cc_data %>%
  mutate(day = get_day(timestamp),
         weekday = weekdays(timestamp),
         hour = get_hour(timestamp),
         time = format(timestamp, format = "%H:00"))

weekday_cc_data <- sel_cc_data %>%
  filter(weekday != "Saturday" |
           weekday != "Sunday")

p1 <- ggplot(weekday_cc_data, 
       aes(factor(weekday_cc_data$time), 
           fill = location)) + 
  scale_fill_brewer(palette = "Set1") + 
  geom_bar(position = position_dodge2(preserve = "single")) +
  geom_text(stat = 'count',
            aes(label = ..count..),
            position = position_dodge2(width = 0.9, preserve = "single"),
            vjust = -0.5,
            size = 3) + 
  ylim(0,105) +
  labs(x = "Time in Hours", y = "Number of Transactions", title = "Weekdays Visits by Time (H)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        plot.title = element_text(hjust = 0.5))

p2 <- ggplot(weekday_cc_data, aes(factor(weekday_cc_data$time), 
                              fill = location)) + 
  scale_fill_brewer(palette = "Set1") + 
  geom_bar() + 
  labs(x = "Time in Hours", y = "Number of Transactions", title = "(Stacked) Weekdays Visits by Time (H)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        plot.title = element_text(hjust = 0.5))

g1 <- ggplotGrob(p1)
g2 <- ggplotGrob(p2)

g <- rbind(g1, g2, size = "first")

g$widths <- unit.pmax(g1$widths, g2$widths)

##########################
#disable plotting
##########################

#grid.newpage()
#grid.draw(g)

weekend_cc_data <- sel_cc_data %>%
  filter(weekday == "Saturday" |
           weekday == "Sunday")

p3 <- ggplot(weekend_cc_data, 
       aes(factor(weekend_cc_data$time), 
           fill = location)) + 
  scale_fill_brewer(palette = "Set1") + 
  geom_bar(position = position_dodge2(preserve = "single")) +
  geom_text(stat = 'count',
            aes(label = ..count..),
            position = position_dodge2(width = 0.9, preserve = "single"),
            vjust = -0.5,
            size = 3) + 
  ylim(0,100) +
  labs(x = "Time in Hours", y = "Number of Transactions", title = "Weekends Visits by Time (H)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        plot.title = element_text(hjust = 0.5))

p4 <- ggplot(weekend_cc_data, aes(factor(weekend_cc_data$time), 
                              fill = location)) + 
  scale_fill_brewer(palette = "Set1") + 
  geom_bar() + 
  labs(x = "Time in Hours", y = "Number of Transactions", title = "(Stacked) Weekends Visits by Time (H)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        plot.title = element_text(hjust = 0.5))

g1 <- ggplotGrob(p1)
g2 <- ggplotGrob(p3)

g <- rbind(g1, g2, size = "first")

g$widths <- unit.pmax(g1$widths, g2$widths)

##########################
#disable plotting
##########################
#grid.newpage()
#grid.draw(g)

```

```{r q2}
#have a count on the popular places
locations_gps <- locations_gps %>%
  group_by(lat, long) %>%
  add_count(lat, long, name = "count")

#discard unnecessary information
locations_gps <- locations_gps %>%
  select(-c(Timestamp, id)) %>%
  distinct(lat, long, count)

#Attempt to filter out locations with more than 100
pop_location_gps <- locations_gps %>%
  filter(count >= 50)

tif_sf <- st_as_sf(pop_location_gps, 
                   coords = c("long", "lat"), 
                   crs = 4326)
tmap_mode("plot")

tm <- tm_shape(bgmap) + 
 tm_rgb(bgmap, r = 1, g = 2, b = 3,
         alpha = NA,
         saturation = 1,
         interpolate = TRUE,
         max.value = 255) +
tm_shape(tif_sf) + 
  tm_dots()
```

```{r gps lag}
gps_lag <- gps %>%
  group_by(id) %>%
  mutate(timelag = lead(Timestamp, order_by = id)) %>%
  mutate(lag = timelag - Timestamp)

y_pp <- format_format(big.mark = "",
                      decimal.mark = ",",
                      scientific = FALSE)

gps_lag_plot <- ggplot(gps_lag, aes(factor(id), lag)) + 
  labs(x = "ID", y = "Time Lag", title = "GPS Lag Interval") +
  scale_y_continuous(labels = y_pp) + 
  geom_boxplot_interactive(aes(tooltip = Timestamp),
                           stackgroups = TRUE) + 
  theme(axis.text.x = element_text(size = 8, angle = 45, vjust = 1.1, hjust = 1.1),
        axis.text.y = element_text(size = 7),
        plot.title = element_text(hjust = 0.5))

#adjust to 10s
adj_gps_lag <- gps_lag %>%
  filter(lag <= 5)

adj_gps_lag_plot <- ggplot(adj_gps_lag, aes(factor(id), lag)) + 
  labs(x = "ID", y = "Time Lag", title = "GPS Lag Interval") +
  scale_y_continuous(labels = y_pp) + 
  geom_boxplot_interactive(aes(tooltip = Timestamp),
                           stackgroups = TRUE) + 
  theme(axis.text.x = element_text(size = 8, angle = 45, vjust = 1.1, hjust = 1.1),
        axis.text.y = element_text(size = 7),
        plot.title = element_text(hjust = 0.5))

POI_gps <- locations_gps %>%
  mutate(min3 = Timestamp + 60*3) %>%
  group_by(id) %>%
  mutate(stay = if_else(min3 <= lag(min3) & lat == lag(lat) & long == lag(long), TRUE, FALSE)) %>% 
  ungroup() %>%
  filter(stay == TRUE) %>%
  distinct(lat, long)

```

## 4.1 Investigating GPS Time Lag

With the gps data, which comes with 5-decimal points for the lat/long co-ordinates, we attempted to round it off to 4 decimal points, to get an accuracy of only 11.1metres, so that we could do a count on popular co-ordinates that was frequented. Knowing that GPS communication would be continuous, it would be good to investigate the time lag between data to data also.

From the boxplot, we could see that there are numerous outliers, such that the actual boxplot is too small to be seen. There are reasons as to why the lag between data is large. The reasons, included but not limited to:

+ The car was parked for a long period of time. Thus, the current to power the GPS could be cut off to prevent the vehicle battery from draining out completely. 

+ GPS is malfunctioning

+ It is programmed not to store any locations if GPS signal is weak.

+ Tampering of GPS

Since we are more interested to infer the GPS data to Point of Interests, we would adjust the data such that we can filter out outliers.

From the adjusted GPS Lag Interval chart, we know that GPS would usually store its location with an interval of 1 to 2 seconds.

## 4.2 Establishing Point Of Interests (POI)

For this set of raw data, it would be useful to establish a set of coordinates that could be inferred to POIs on the map. To do so, we would examine every set of GPS data and compare it to its location **3 mins later**. If its coordinates are the same, the assumption would be that the vehicle is parked at a location which may be inferred to as POI. And for the identification of POIs, we do not need a set of GPS coordinates that is highly accurate. As such, we would use a set of GPS coordinates that is accurate to the nearest 11.1m.

```{r 10000 transaction, include=FALSE}

txn_time <- date_time_parse("2014-01-13 19:19:00",
                zone = "",
                format = "%Y-%m-%d %H:%M")

POI_indi <- locations_gps %>%
  mutate(Datestamp = as.Date(Timestamp + 60*60*8)) %>%
  filter(Datestamp == txn_time) %>%
  mutate(stop = Timestamp - lag(Timestamp)) %>%
  mutate(parked = ifelse(stop >60*3, TRUE,FALSE)) %>%
  filter(parked == TRUE) %>%
  filter(lat == 24.86 & long == 36.07) %>%
  distinct(lat, long)
    
  tif_sf_indi <- st_as_sf(POI_indi, 
                   coords = c("long", "lat"), 
                   crs = 4326)

  tm <- tmPOI + 
    tm_shape(tif_sf_indi) + 
    tm_dots(size = 0.8,
            alpha = 0.3, 
            col = "green")

pir <- paste("Route of Fryos ", " on ", as.Date(end_time + 60*60*8),".png", sep = "")
tmap_save(plot_Fryros, filename = pir)

```

```{r display txn10000 records, echo=FALSE}
txn10000 <- txn10000[order(txn10000$Timestamp),]

txn10000 %>%
  kbl() %>%
  kable_styling()
```


```{r print 33, echo=FALSE}
plot_txn10000
```

```{r location records for 9551, echo=FALSE}
#finding locations with a 2 min buffer around 13/1/14 19:20
start_time <- date_time_parse("2014-01-13 00:00:00",
                zone = "",
                format = "%Y-%m-%d %H:%M")

end_time <- date_time_parse("2014-01-13 23:59:00",
                zone = "",
                format = "%Y-%m-%d %H:%M")

txn9551 <- cc_data %>%
  filter(last4ccnum == 9551) %>%
  filter(timestamp > start_time & timestamp < end_time) %>%
  select(timestamp, location, price)

txn9551 %>%
  kbl() %>%
  kable_styling()
```

```{r 2276 txn, echo=FALSE}

txn2276 <- records_cc %>%
  filter(last4ccnum == 2276 & Datestamp == "2014-01-14") %>%
  select(-c(Datestamp, last4ccnum, "Visits on Day")) %>%
  kbl() %>%
  kable_styling()

txn2276

places14 <- records_POI %>%
  filter(Datestamp == "2014-01-14" & "Visits on Day" == 3)

```